---
title: "Movie Recommendation System"
author: "Aviparna Biswas"
output: html_notebook
---

### 0. Dataset

In order to build our recommendation system, we have used the MovieLens Dataset. It consists of two files movies.csv and ratings.csv. This data consists of 105339 ratings applied over 10329 movies.

### 1. Importing Essential Libraries

```{r}

library(recommenderlab)

```

```{r}
library(ggplot2)
library(data.table)
library(reshape2)
```

### 2. Retrieving the Data

We will now retrieve our data from movies.csv into movie_data dataframe and ratings.csv into rating_data. We will use the str() function to display information about the movie_data dataframe.

```{r}
movie_data <- read.csv("movies.csv",stringsAsFactors=FALSE)
rating_data <- read.csv("ratings.csv")
str(movie_data)
```

#### 2.1. Checking the movie_data
```{r}
summary(movie_data)
```
```{r}
head(movie_data)
```
#### 2.2. Checking the rating_data
```{r}
summary(rating_data)
```
```{r}
head(rating_data)
```
### 3. Data Pre-processing
From the above table, we observe that the userId column, as well as the movieId column, consist of integers. Furthermore, we need to convert the genres present in the movie_data dataframe into a more usable format by the users. In order to do so, we will first create a one-hot encoding to create a matrix that comprises of corresponding genres for each of the films.

```{r}
movie_genre <- as.data.frame(movie_data$genres, stringsAsFactors = FALSE)
library(data.table)

movie_genre2 <- as.data.frame(tstrsplit(movie_genre[,1], '[|]', type.convert = TRUE), stringsAsFactors = FALSE)

colnames(movie_genre2) <- c(1:10)

list_genre <- c("Action", "Adventure", "Animation", "Children", "Comedy", "Crime","Documentary", "Drama", "Fantasy", "Film-Noir", "Horror", "Musical", "Mystery","Romance", "Sci-Fi", "Thriller", "War", "Western")

genre_mat1 <- matrix(0, 10330, 18)
genre_mat1[1,] <- list_genre
colnames(genre_mat1) <- list_genre

for (index in 1:nrow(movie_genre2)) {
  for (col in 1:ncol(movie_genre2)) {
    gen_col = which(genre_mat1[1,] == movie_genre2[index, col])
    genre_mat1[index+1, gen_col] <- 1
  }
}

genre_mat2 <- as.data.frame(genre_mat1[-1,], stringsAsFactors = FALSE)

for (col in 1:ncol(genre_mat2)) {
  genre_mat2[,col] <-  as.integer(genre_mat2[,col])
}

str(genre_mat2)
```
#### 3.1. Creating a Search Matrix 
```{r}
SearchMatrix <- cbind(movie_data[,1:2], genre_mat2[])
head(SearchMatrix)
```
#### 3.2. Creating a sparse matrix
```{r}
ratingMatrix <- dcast(rating_data, userId~movieId, value.var = "rating", na.rm=FALSE)
ratingMatrix <- as.matrix(ratingMatrix[,-1])

ratingMatrix <- as(ratingMatrix, "realRatingMatrix")
ratingMatrix
```

```{r}
recommendation_model <- recommenderRegistry$get_entries(dataType = "realRatingMatrix")
names(recommendation_model)
```
```{r}
lapply(recommendation_model, "[[", "description")
```
```{r}
recommendation_model$IBCF_realRatingMatrix$parameters
```
### 4. Exploring Similar Data
Collaborative Filtering involves suggesting movies to the users that are based on collecting preferences from many other users. For example, if a user A likes to watch action films and so does user B, then the movies that the user B will watch in the future will be recommended to A and vice-versa. Therefore, recommending movies is dependent on creating a relationship of similarity between the two users. With the help of recommenderlab, we can compute similarities using various operators like cosine, pearson as well as jaccard.
```{r}
similarity_mat <- similarity(ratingMatrix[1:4, ], method = "cosine", which = "users")
as.matrix(similarity_mat)

image(as.matrix(similarity_mat), main = "User's Similarities")
```
In the above matrix, each row and column represents a user. We have taken four users and each cell in this matrix represents the similarity that is shared between the two users.

Now, we delineate the similarity that is shared between the films:

```{r}
movie_similarity <- similarity(ratingMatrix[, 1:4], method = "cosine", which = "items")
as.matrix(movie_similarity)

image(as.matrix(movie_similarity), main = "Movies similarity")
```
#### 4.1. Extracting the unique ratings
```{r}
rating_values <- as.vector(ratingMatrix@data)
unique(rating_values)
```
#### 4.2. Table with most unique ratings
```{r}
Table_of_Ratings <- table(rating_values)
Table_of_Ratings
```
### 5. Most viewed movies visualization
```{r}
library(ggplot2)
movie_views <- colCounts(ratingMatrix)
table_views <- data.frame(movie = names(movie_views), views = movie_views)

table_views <- table_views[order(table_views$views, decreasing = TRUE), ]
table_views$title <- NA

for (index in 1:10325) {
  table_views[index, 3] <- as.character(subset(movie_data, movie_data$movieId == table_views[index, 1])$title)
}
table_views[1:6,]
```
#### 5.1. Visualize a bar plot for the total number of views of the top films
```{r}
ggplot(table_views[1:6, ], aes(x = title, y = views)) +
  geom_bar(stat="identity", fill = 'steelblue') +
  geom_text(aes(label=views), vjust=-0.3, size=3.5) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +

  ggtitle("Total Views of the Top Films")
```

