---
title: "Movie Recommendation System"
author: "Aviparna Biswas"
date: "28-05-2022"
output: html_notebook
---

### 0. Dataset

In order to build our recommendation system, we have used the MovieLens Dataset. It consists of two files movies.csv and ratings.csv. This data consists of 105339 ratings applied over 10329 movies.

### 1. Importing Essential Libraries

```{r}

library(recommenderlab)

```

```{r}
library(ggplot2)
library(data.table)
library(reshape2)
```

### 2. Retrieving the Data

We will now retrieve our data from movies.csv into movie_data dataframe and ratings.csv into rating_data. We will use the str() function to display information about the movie_data dataframe.

```{r}
movie_data <- read.csv("movies.csv",stringsAsFactors=FALSE)
rating_data <- read.csv("ratings.csv")
str(movie_data)
```

#### 2.1. Checking the movie_data
```{r}
summary(movie_data)
```
```{r}
head(movie_data)
```
#### 2.2. Checking the rating_data
```{r}
summary(rating_data)
```
```{r}
head(rating_data)
```
### 3. Data Pre-processing
From the above table, we observe that the userId column, as well as the movieId column, consist of integers. Furthermore, we need to convert the genres present in the movie_data dataframe into a more usable format by the users. In order to do so, we will first create a one-hot encoding to create a matrix that comprises of corresponding genres for each of the films.

```{r}
movie_genre <- as.data.frame(movie_data$genres, stringsAsFactors = FALSE)
library(data.table)

movie_genre2 <- as.data.frame(tstrsplit(movie_genre[,1], '[|]', type.convert = TRUE), stringsAsFactors = FALSE)

colnames(movie_genre2) <- c(1:10)

list_genre <- c("Action", "Adventure", "Animation", "Children", "Comedy", "Crime","Documentary", "Drama", "Fantasy", "Film-Noir", "Horror", "Musical", "Mystery","Romance", "Sci-Fi", "Thriller", "War", "Western")

genre_mat1 <- matrix(0, 10330, 18)
genre_mat1[1,] <- list_genre
colnames(genre_mat1) <- list_genre

for (index in 1:nrow(movie_genre2)) {
  for (col in 1:ncol(movie_genre2)) {
    gen_col = which(genre_mat1[1,] == movie_genre2[index, col])
    genre_mat1[index+1, gen_col] <- 1
  }
}

genre_mat2 <- as.data.frame(genre_mat1[-1,], stringsAsFactors = FALSE)

for (col in 1:ncol(genre_mat2)) {
  genre_mat2[,col] <-  as.integer(genre_mat2[,col])
}

str(genre_mat2)
```
#### 3.1. Creating a Search Matrix 
```{r}
SearchMatrix <- cbind(movie_data[,1:2], genre_mat2[])
head(SearchMatrix)
```
#### 3.2. Creating a sparse matrix
```{r}
ratingMatrix <- dcast(rating_data, userId~movieId, value.var = "rating", na.rm=FALSE)
ratingMatrix <- as.matrix(ratingMatrix[,-1])

ratingMatrix <- as(ratingMatrix, "realRatingMatrix")
ratingMatrix
```

```{r}
recommendation_model <- recommenderRegistry$get_entries(dataType = "realRatingMatrix")
names(recommendation_model)
```
```{r}
lapply(recommendation_model, "[[", "description")
```
```{r}
recommendation_model$IBCF_realRatingMatrix$parameters
```
### 4. Exploring Similar Data
Collaborative Filtering involves suggesting movies to the users that are based on collecting preferences from many other users. For example, if a user A likes to watch action films and so does user B, then the movies that the user B will watch in the future will be recommended to A and vice-versa. Therefore, recommending movies is dependent on creating a relationship of similarity between the two users. With the help of recommenderlab, we can compute similarities using various operators like cosine, pearson as well as jaccard.
```{r}
similarity_mat <- similarity(ratingMatrix[1:4, ], method = "cosine", which = "users")
as.matrix(similarity_mat)

image(as.matrix(similarity_mat), main = "User's Similarities")
```
In the above matrix, each row and column represents a user. We have taken four users and each cell in this matrix represents the similarity that is shared between the two users.

Now, we delineate the similarity that is shared between the films:

```{r}
movie_similarity <- similarity(ratingMatrix[, 1:4], method = "cosine", which = "items")
as.matrix(movie_similarity)

image(as.matrix(movie_similarity), main = "Movies similarity")
```
#### 4.1. Extracting the unique ratings
```{r}
rating_values <- as.vector(ratingMatrix@data)
unique(rating_values)
```
#### 4.2. Table with most unique ratings
```{r}
Table_of_Ratings <- table(rating_values)
Table_of_Ratings
```
### 5. Most viewed movies visualization
```{r}
library(ggplot2)
movie_views <- colCounts(ratingMatrix)
table_views <- data.frame(movie = names(movie_views), views = movie_views)

table_views <- table_views[order(table_views$views, decreasing = TRUE), ]
table_views$title <- NA

for (index in 1:10325) {
  table_views[index, 3] <- as.character(subset(movie_data, movie_data$movieId == table_views[index, 1])$title)
}
table_views[1:6,]
```
#### 5.1. Visualize a bar plot for the total number of views of the top films
```{r}
ggplot(table_views[1:6, ], aes(x = title, y = views)) +
  geom_bar(stat="identity", fill = 'steelblue') +
  geom_text(aes(label=views), vjust=-0.3, size=3.5) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +

  ggtitle("Total Views of the Top Films")
```
### 6. Heatmap of Movie Ratings
```{r}
image(ratingMatrix[1:20, 1:25], axes = FALSE, main = "Heatmap of the first 25 rows and 25 columns")
```
### 7. Performing Data Preparation

#### 7.1. Selecting the useful data
For finding useful data in our dataset, we have set the threshold for the minimum number of users who have rated a film as 50. This is also same for minimum number of views that are per film. This way, we have filtered a list of watched films from least-watched ones.
```{r}
movie_ratings <- ratingMatrix[rowCounts(ratingMatrix) > 50, colCounts(ratingMatrix) > 50]

movie_ratings
```
From the above output of ‘movie_ratings’, we observe that there are 420 users and 447 films as opposed to the previous 668 users and 10325 films. We can now delineate our matrix of relevant users as follows:
```{r}
minimum_movies<- quantile(rowCounts(movie_ratings), 0.98)
minimum_users <- quantile(colCounts(movie_ratings), 0.98)
image(movie_ratings[rowCounts(movie_ratings) > minimum_movies,
                     colCounts(movie_ratings) > minimum_users],
main = "Heatmap of the top users and movies")
```
Now, we will visualize the distribution of the average ratings per user.
```{r}
average_ratings <- rowMeans(movie_ratings)
qplot(average_ratings, fill=I("steelblue"), col=I("red")) +
  ggtitle("Distribution of the average rating per user")
```
#### 7.2. Data Normalization
Normalization is a data preparation procedure to standardize the numerical values in a column to a common scale value. This is done in such a way that there is no distortion in the range of values. Normalization transforms the average value of our ratings column to 0. We then plot a heatmap that delineates our normalized ratings.
```{r}
normalized_ratings <- normalize(movie_ratings)
sum(rowMeans(normalized_ratings) > 0.0001)

image(normalized_ratings[rowCounts(normalized_ratings) > minimum_movies, colCounts(normalized_ratings) > minimum_users], main = "Normalized Ratings of the Top Users")
```
#### 7.3. Data Binarization
Binarizing the data means that we have two discrete values 1 and 0, which will allow our recommendation systems to work more efficiently. We will define a matrix that will consist of 1 if the rating is above 3 and otherwise it will be 0.
```{r}
binary_minimum_movies <- quantile(rowCounts(movie_ratings), 0.95)
binary_minimum_users <- quantile(colCounts(movie_ratings), 0.95)
#movies_watched <- binarize(movie_ratings, minRating = 1)

good_rated_films <- binarize(movie_ratings, minRating = 3)
image(good_rated_films[rowCounts(movie_ratings) > binary_minimum_movies,
colCounts(movie_ratings) > binary_minimum_users],
main = "Heatmap of the top users and movies")
```
### 8. Collaborative Filtering System
The collaborative filtering finds similarity in the items based on the people’s ratings of them. The algorithm first builds a similar-items table of the customers who have purchased them into a combination of similar items. This is then fed into the recommendation system.
```{r}
sampled_data<- sample(x = c(TRUE, FALSE),
                      size = nrow(movie_ratings),
                      replace = TRUE,
                      prob = c(0.8, 0.2))
training_data <- movie_ratings[sampled_data, ]
testing_data <- movie_ratings[!sampled_data, ]
```
### 9. Building the Recommendation System
We will now explore the various parameters of our Item Based Collaborative Filter. These parameters are default in nature. In the first step, k denotes the number of items for computing their similarities. Here, k is equal to 30. Therefore, the algorithm will now identify the k most similar items and store their number. We use the cosine method which is the default one but you can also use pearson method.
```{r}
recommendation_system <- recommenderRegistry$get_entries(dataType ="realRatingMatrix")
recommendation_system$IBCF_realRatingMatrix$parameters
```
```{r}
recommen_model <- Recommender(data = training_data,
                          method = "IBCF",
                          parameter = list(k = 30))
recommen_model
class(recommen_model)
```
```{r}
model_info <- getModel(recommen_model)
class(model_info$sim)
dim(model_info$sim)
top_items <- 20
image(model_info$sim[1:top_items, 1:top_items],
   main = "Heatmap of the first rows and columns")
```
```{r}
sum_rows <- rowSums(model_info$sim > 0)
table(sum_rows)

sum_cols <- colSums(model_info$sim > 0)
qplot(sum_cols, fill=I("steelblue"), col=I("red"))+ ggtitle("Distribution of the column count")
```
### 10. Recommender System on the dataset
We will create a top_recommendations variable which will be initialized to 10, specifying the number of films to each user. We will then use the predict() function that will identify similar items and will rank them appropriately. Here, each rating is used as a weight. Each weight is multiplied with related similarities. Finally, everything is added in the end.
```{r}
top_recommendations <- 10 # the number of items to recommend to each user
predicted_recommendations <- predict(object = recommen_model,
                          newdata = testing_data,
                          n = top_recommendations)
predicted_recommendations
```
```{r}
user1 <- predicted_recommendations@items[[1]] # recommendation for the first user
movies_user1 <- predicted_recommendations@itemLabels[user1]
movies_user2 <- movies_user1
for (index in 1:10){
  movies_user2[index] <- as.character(subset(movie_data,
                                         movie_data$movieId == movies_user1[index])$title)
}
movies_user2
```
```{r}
recommendation_matrix <- sapply(predicted_recommendations@items,
                      function(x){ as.integer(colnames(movie_ratings)[x]) }) # matrix with the recommendations for each user
#dim(recc_matrix)
recommendation_matrix[,1:4]
```

